\documentclass[12pt]{article}

\include{../style}

\begin{document}

\input{title}

\textbf{Credit Statement:} Talked to Sair Shaikh'26, and Math Stack Exchange.
\\

\begin{problem}{1}
    \textit{Solvable up to sixty!} Recall that $A_5$, which has order $60$, is simple and nonabelian. The goal is to:
    \begin{enumerate}
        \item[($\aleph$)] Prove that all groups of order $< 60$ are solvable.
    \end{enumerate}
    
    This has two steps. 
    \begin{enumerate}
        \item [($\alpha$)] First, use Jordan–Hölder to prove that if $60$ is the first order of a finite nonabelian simple group, then all groups of order $< 60$ are solvable.
        \item [($\beta$)] Second, prove that every nonabelian group of order $< 60$ is not simple. Since the abelian simple groups are precisely those of prime order, for each composite order $n < 60$, we will try to prove that any group of order $n$ is not simple. For example, we already know that no group of order $p^\alpha$, with $\alpha > 1$, is simple and that no group of order $pq$, with $p$ and $q$ primes, is simple. Prove the following additional criteria on the order of a group for the group to not be simple:
    \end{enumerate}

    \textbf{Hints.} 
   Part (a) follows from a direct application of the congruence conditions in the Sylow theorems. For (b), assume the contrary and consider the possible number of Sylow $r$-subgroups; then use this to count the number of elements of order $r$ (any two Sylow $r$-subgroups intersect only at the identity). Combine this with the number of elements of order $p$ and $q$ to find more elements than the order of the group. For (c) and (d), handle small $k$ using the Sylow congruence conditions, and then for large $k$, consider the permutation representation associated to the conjugation action of $G$ on the set of Sylow $2$-subgroups. For (e) and (f), do the same using the Sylow $3$-subgroups. For (g), if neither the Sylow $2$- nor $7$-subgroups are normal, start counting elements in these subgroups to reach a contradiction (while any two Sylow $7$-subgroups only intersect at the identity, how could Sylow $2$-subgroups intersect?).
    
    Finally, use all the criteria you know to handle every composite order $< 60$. Have fun! How much higher can you go using these same tools?
\end{problem}

\newpage

\begin{problem} {($\alpha$)}
    First, use Jordan–Hölder to prove that if 60 is the first order of a finite nonabelian simple group, then all groups of order $<$ 60 are solvable.
\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    If 60 is the first order of a finite non-abelian simple group, then groups of $< 60$ are either abelian, or simple (they're all finite). 
    \begin{enumerate}
        \item \textit{The group is abelian.} By Jordan–Hölder we know that a composition series for group $G$ exists. Because $G$ is abelian, all subgroups are normal $\implies$ all composition factors are normal $\implies$ group is solvable. 
        \item \textit{The group is non-abelian and non-simple.} By Jordan–Hölder we know that a composition series for group $G$ exists. Assume all groups of order less than $n$ are solvable, for some $n < 59$. Then because group $G$ of order $n$ is not simple, we know that at least one non trivial normal subgroup $N$ exists. Obviously, the orders of $N$ and $G/N$ are both less than $n$. Then by our induction hypothesis we know that both $N$ and $G/N$ are solvable, and hence $G$ is solvable. 
    \end{enumerate}
\end{solution}

\newpage

\begin{problem}{($\beta .a$)}

    If $G$ is a finite group of order $p^{k}m$, with $p \nmid m$ and $m < p$ (more generally, no divisor of $m$ other than $1$ is congruent to $1$ modulo $p$), then $G$ has a normal Sylow $p$-subgroup.

\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    By Sylow's Theorem, we know that $G$ ha a Sylow $p$-subgroup $P$ of order $p^k$. We also know that the following to relations must hold: 
    \[n_p \mid m ,\qquad n_p \equiv 1 \text{ (mod $p$)}\]
    where $n_p$ is the number of Sylow p-groups in $G$. But the only divisor of $m$ congruent to 1 modulo $p$ is 1 (given). Hence $n_p = 1$, and the Sylow $p$-group is unique. \\
    And by Corollary 20 if $P$ is the unique Sylow $p$-group in $G$, then it is also normal in $G$. Hence $G$ has a normal Sylow $p-$subgroup and $G$ is not simple. 
\end{solution}

\newpage 

\begin{problem}{($\beta .$b)}

    If $G$ is a finite group of order $pqr$, where $p, q, r$ are primes with $p < q < r$, then $G$ has a normal Sylow subgroup for at least one of $p, q,$ or $r$.

\end{problem}

\begin{solution}
    \bbni 
    \bbni
    By Sylow's Theorems, we know the following, 
    \begin{enumerate}
        \item $G$ has at least one Sylow $r$-subgroup. 
        \item The number of Sylow $r$-subgroups $n_r$ divides $pq$, and $n_r \equiv 1 \text{ (mod $r$)}$. 
    \end{enumerate}
    Then,  $n_r \mid pq\implies  n_r = 1, p , q, pq$. But $n_r \neq p, q$ because both $p$ and $q$ are less than $r$ and hence not congruent to 1 mod $r$. I.e. $n_r = 1, pq$. \\
    If $n_r = 1$, then by part (a), $G$ has a normal Sylow $r$-subgroup. If $n_r = pq$, then there are $pq$ Sylow $r$-subgroups. 
    Because $r$ is prime, all $pq$ subgroups are cyclic, and intersect only at identity. I.e. there are $pq (r -1)$ elements of order $r$ in $G$, and $pq$ elements of other orders. 

    Similarly, we know, 
     \begin{enumerate}
        \item $G$ has at least one Sylow $q$-subgroup. 
        \item The number of Sylow $q$-subgroups $n_r$ divides $pr$, and $n_q \equiv 1 \text{ (mod $q$)}$. 
    \end{enumerate}
    Then,  $n_q \mid pr\implies  n_q = 1, p , r, pr$. But $n_q \neq p$ because $p$ is a prime number less than $q$ and hence not congruent to 1 mod $q$. 
    
    I.e. $n_q = 1, r, pr$. \\
    If $n_q = 1$, then by part (a), $G$ has a normal Sylow $q$-subgroup. 
    If $n_q = r$, then there are $r$ Sylow $q$-subgroups. And because $q$ is prime, all $r$ subgroups are cyclic, and interesect only at identity. I.e. there are $r (q -1)$ elements of order $q$ in $G$. 
    
    If $n_q = pr$, then there are $pr$ Sylow $q$-subgroups. 
    Because $q$ is prime, all $pr$ subgroups are cyclic, and intersect only at identity. I.e. there are $pr (q -1)$ elements of order $q$ in $G$. 

    If we assume that $n_r \neq 1$ and $n_q \neq 1$, then we have 2 cases, 
    \begin{enumerate}
        \item $n_r = pq, n_q =r$
             \begin{align*}
                    \implies pq &\geq rq - r \\
                    pq &\geq r(q-1)
                \end{align*}
            But $r > q$ and $q-1 > p$, so this is a contradiction and at least one of $n_r$ or $n_q$ needs to be 1.  
        \item $n_r = pq, n_q = pr$
             \begin{align*}
                \implies pq &\geq pqr - pr \\
                q &\geq qr -r \\
                q &\geq r (q - 1)
            \end{align*}
            But $r > q$ and $q-1 > 1$, so this is a contradiction and at least one of $n_r$ or $n_q$ needs to be 1.  
    \end{enumerate}

    Thus, there exists at least 1 normal Sylow $p$-subgroup in $G$, and $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta .c$)}

    If $G$ is a finite group of order $2^k \cdot 3$, with $k \ge 1$, then $G$ is not simple.

\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    Consider the Sylow $2$-group of order $2^k$, then $n_2 = 1, 3$. If $n_2 = 1$, then the Sylow subgroup is unique and by Corollary 20 also normal. If $n_2 = 3$, then by Sylow's Second Theorem all 3-Sylow subgroups are conjugate to each other. Let the set of these subgroups be $A = \{P_1, P_2, P_3\}$, then we can represent the conjugation action of $G$ on $A$ by $\{g \in G: gP_ig^{-1} \in S\}$. The permutation representation of this action can be defined as $\varphi: G \to S_{\mid A \mid}$. Then we have the following cases, 
    \begin{enumerate}
        \item If the kernel is not trivial, then $\varphi$ is not injective, and $G$ has a normal subgroup. For $\varphi$ to not be trivial, $| G | \, > \, | S_A  | \implies | G |  \, > 6 \implies k> 1$.
        \item If the kernel is trivial, or $k = 1$, then by part (a) $G$ has a normal subgroup. 
    \end{enumerate}
    Hence, $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta .d$)}

    If $G$ is a finite group of order $2^k \cdot 5$, with $k \ge 1$, then $G$ is not simple.

\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    Similar to the argument above, consider the Sylow $2$-group of order $2^k$, then $n_2 = 1, 5$. If $n_2 = 1$, then the Sylow subgroup is unique and by Corollary 20 also normal. If $n_2 = 5$, then by Sylow's Second Theorem all 5-Sylow subgroups are conjugate to each other. Let the set of these subgroups be $A = \{P_1, P_2, P_3, P_4, P_5\}$, then we can represent the conjugation action of $G$ on $A$ by $\{g \in G: gP_ig^{-1} \in S\}$. The permutation representation of this action can be defined as $\varphi: G \to S_{\mid A \mid}$. Then we have the following cases, 
    \begin{enumerate}
        \item If the kernel is not trivial, then $\varphi$ is not injective, and $G$ has a normal subgroup. For $\varphi$ to not be trivial, $| G | \, > \, | S_A  | \implies | G |  \, > 120 \implies k> 4$.
        \item If the kernel is trivial, or $k = 1$, then by part (a) $G$ has a normal subgroup. 
        \item If $k = 2$, then by part (a) $G$ has a normal subgroup. 
        \item If $k = 3$, then the number of 5-Sylow groups $n_5 \mid 8 \implies n_5 = 1, 2, 4, 8$ but only $1 \equiv 1$ (mod $5$). 
        \item If $k=4$, then $\mid G \mid = 80$, and $ 80 \nmid 120$, so $\not \exists$ a subgroup of $S_5$ that $G$ is isomorphic to. I.e $\varphi$ is not injective, has a non-trivial kernel, and hence also a normal subgroup. 
    \end{enumerate}
    Hence, $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta .e$)}

    If $G$ is a finite group of order $2^2 \cdot 3^k$, with $k \ge 1$, then $G$ is not simple. For $k = 1$, use part (c).
    
\end{problem}

\begin{solution}
    \bbni
    \bbni
    There are 2 cases, if $k=1$, then $G$ is a finite group of order $12$, and can be represented as $|G| = 2^k\cdot 3$, where $k = 1$. Hence $G$ is not simple by part (c). If $k \geq 2$, then we know  by Sylow's First Theorem that there exists at least one Sylow 3-subgroup of order $3^k$ in $G$. The number of Sylow 3-subgroups $n_3$ divides 4, i.e. $n_3 = 1, 2, 4$. But $n_3$ is also congruent to 1 mod 3, so $n_3$ must be 1. By Corollory 20, we know that if the Sylow p-subgroup is unique it is also normal. Hence, $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta .f$)}

    If $G$ is a finite group of order $3^k \cdot 5$, with $k \ge 1$, then $G$ is not simple.

\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    We know  by Sylow's First Theorem that there exists at least one Sylow 3-subgroup of order $3^k$ in $G$. The number of Sylow 3-subgroups $n_3$ divides 5, i.e. $n_3 = 1, 5$. But $n_3$ is also congruent to 1 mod 3, so $n_3$ must be 1. By Corollory 20, we know that if the Sylow p-subgroup is unique it is also normal. Hence, $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta . g$)}

    No group of order $56$ is simple.    

\end{problem}

\begin{solution}
    \bbni 
    \bbni 
    The prime factorization of 56 is $2 ^3 \cdot 7$. We know  by Sylow's First Theorem that there exists at least one Sylow 7-subgroup of order $7$ in $G$. The number of Sylow 7-subgroups $n_7$ divides 8, i.e. $n_7 = 1, 2, 4, 8$. But $n_7$ is also congruent to 1 mod 7, so $n_7$ can only be either 1 or 8. If $n_7 = 1$ then by Corollory 20, we know that if a Sylow p-group is unique it is also normal, then we are done. But if $n_7 = 8$, then we know that the conjugates intersect only at identity (cyclic groups), i.e. the number of elements of order 7 (all elements in a cyclic group of prime order have order equal to the prime) is $8 ( 7-1) = 42$. So there exist 8 elements of order not equal to 7. 
    
    Again by Sylow's First Theorem we also know that there exists at least one Sylow 2-subgroup of order $8$ in $G$. The order of none of these elements is 7, because $7 \nmid 8$. Then we can say that there only exists 1 Sylow 2-subgroup. And by Corollory 20, we know that if the Sylow p-subgroup is unique it is also normal. 
    Hence, $G$ is not simple. 
\end{solution}

\newpage

\begin{problem}{($\beta . h$)}

    \textit{Optional.} What is the largest number $N > 60$ you can find such that no group of composite order $n$, with $61 \le n \le N$, is simple?
    
\end{problem}

\begin{solution}
    \bbni 
    \bbni
    Using just the facts we have proved so far, we can show that groups with order upto $71$ are not simple. But if we prove that all groups of order 72 are not simple, then we can go upto 83. 

    \textit{Proof for $| G | = 72$}. \\
     $| G | = 72 = 2^3 \cdot 3^2$. But for this group consider the Sylow 3-subgroup, this subgroup has order 9, and the number of Sylow 3-subgroups $n_3 \mid 8 \implies n_3 = 1, 2, 4, 8$, but $n_3$ is also congruent to 1 mod 3 $\implies n_3 = 1, 4$. If $n_3 = 1$, then a normal subgroup exists and $G$ is not simple. If $n_3 = 4$ then we can define a homomorphism from $\varphi : G \to S_4$, but $| S_4 | < | G|$, i.e $\varphi$ is not injective, i.e the kernel is not trivial, i.e. a normal subgroup exists. 
    Hence, the group of order 72 is not simple. 
\end{solution}



\end{document}