\documentclass[12pt]{article}

\include{../style}

\begin{document}

\input{title}

\begin{problem}{1}
Prove that any positive integer $n$ greater than $1$ can be written as a product of primes.
\end{problem}

\begin{solution}
    \bbni 

    Let $P(n) = \forall n \in \Z_{>1}: \text{n is a product of primes}$. 

    Assume that $P(k)$ holds where $ 1 < k < n$. I.e. the inductive hypothesis is that any integer greater than $1$ and less than $n$ can be written as a product of primes. 

    Then we have two cases: 
    \begin{enumerate}
        \item $n$ is prime: 

        Then $n$ is a product of itself, and we are done. 

        \item $n$ is not prime: 
        
        Then we can write $n$ as a product of two integers $a, b$. Then, 
        \begin{enumerate}
            \item $a,b$ are prime. 
            
            By definition, $n = a * b$. Hence, $n$ can be written as a product of primes. 

            \item Either one or both of $a,b$ is not prime. 
            
            Since $ b = \frac{n}{a}, a = \frac{n}{b}$ we can see that  $a, b < n$. Then by the induction hypothesis, both $a$ and $b$ can be written as products of primes. Therefore, their product $n = ab$ can also be written as a product of primes.

        \end{enumerate}   
        
    \end{enumerate}
\end{solution}

\newpage 

\begin{problem}{2}
Prove that for every real number $r \neq 1$ and every nonnegative integer $n$,
\[
\sum_{i=0}^{n} r^i = \frac{r^{n+1}-1}{r-1}.
\]
\end{problem}

\begin{solution}
    \bbni 

    Let $P(x)$ be the statement:
\[
\sum_{i=0}^{x} r^i = \frac{r^{x+1}-1}{r-1},
\]
for a fixed real number $r \neq 1$.

    Assume $P(x)$ holds upto $x = n -1$. Then we have two cases: $x =0$ or $x > 0$ 
    
    \begin{enumerate}
        \item Base Case: $x = 0$
    
        $\sum_{i=0}^{0} r^i = r^0 = 1$ 
        and  $\frac{r^{1} - 1}{r-1} = \frac{r-1}{r-1} = 1$. Hence, $P(0)$ holds. 

        \item $x = n$, then $P(n)$: 
        \begin{align*}
        \sum_{i=0}^{n} r^i &= \sum_{i=0}^{n-1} r^i + r^n \\ 
        &= \frac{r^{n}-1}{r-1} + r^n \\ 
        &= \frac{r^{n}-1}{r-1} + \frac{r^n(r -1)}{r -1} \\
        &= \frac{r^{n}-1 +r^{n+1} - r^n}{r -1} \\
        &= \frac{r^{n+1} -1 }{r -1} \\
    \end{align*} 

    \end{enumerate}
    Hence the proposition holds. 
\end{solution}

\newpage

\begin{problem}{3}
    Prove that, given an unlimited supply of $6$-cent coins, $10$-cent coins, and $15$-cent coins, one can make change for any amount that is at least $30$ cents. 
 
    \emph{[Hint: You need many cases.]}
\end{problem}

\begin{solution}
    \bbni

    Let $P(n)= \forall n \geq 30: n$ cents can be made using $6$, $10$, and $15$-cent coins. 

    Assume that for any integer $x$ such that $30 \leq x < n$, we can make $x$ cents using $6, 10, 15$-cent coins. 
    Then there are seven cases to consider: $x = 30, x = 31, x = 32, x = 33, x = 34, x = 35, x > 35$
    \begin{enumerate}
        \item $30$ = $10  + 10 + 10$   
        \item $31$ = $10  + 15 + 6$  
        \item $32$ = $10  + 10 + 6 + 6$  
        \item $33$ = $15  + 6 + 6 + 6$  
        \item $34$ = $10  + 6 + 6 + 6 + 6$  
        \item $35$ = $10  + 10 + 15$  
        \item $x > 35$
        
        If $n > 35$, then $n-6 \geq 30$. By the induction hypothesis, $n-6$ cents can be made. Adding one $6$-cent coin gives us a representation for $n$ cents.
    \end{enumerate} 
    Therefore for any $n \geq 30$, we can make $n$ cents using $6, 10, 15$-cent coins. 
\end{solution}
\end{document}