\documentclass[12pt]{article}

\include{../style}
\usepackage{tikz}
\usetikzlibrary{calc} 

\begin{document}

\input{title}
\begin{problem}{1}
Archimedes discovered a construction that trisects any given angle using a compass and a “marked” straightedge. This straightedge has two special markings a distance $1$ apart.
The straightedge can be placed so that it passes through any already drawn point and such that both markings intersect other already drawn lines or circles.
\begin{center}
\begin{tikzpicture}[scale=2.75,thick]
\draw (-1.4cm,0) -- (2.3cm,0);
\draw (1cm,0) arc (0:180:1cm);
\draw (0,0) -- (120:1.4);
\draw (${2*cos(20)}*(1,0)$) -- +(160:3cm);
\draw (0,0) node[anchor=south east,xshift=-2mm,yshift=0mm] {$\theta$};
\draw (${2*cos(20)}*(1,0)$) node[anchor=south east,xshift=-10mm,yshift=-.8mm] {$\phi$};
\draw (120:1) node[anchor=north,xshift=-1mm,yshift=0mm] {$A$};
\draw[fill=black] (${2*cos(20)}*(1,0)$) circle (0.8pt);
\draw[fill=black] (20:1cm) circle (0.8pt);
\end{tikzpicture}
\end{center}

In the picture, start with a horizontal line and another line meeting it with angle $\theta$. Draw a circle of radius $1$ around the intersection point of these two lines. The other line meets the circle at point $A$. Then place the marked straightedge so that it passes through the point $A$ and such that the markings intersect the circle and the horizontal line. Prove that the angle $\varphi$ that the straightedge makes with the horizontal line is equal to $\theta/3$.
\end{problem}

\begin{problem}{2}
Prove that any angle $\theta$ such that $\tan(\theta)$ is rational can be constructed with compass and straightedge. Prove that such angles are dense in the interval $(-\pi/2,\pi/2)$.
\end{problem}

\begin{problem}{3}
Prove that an angle $\theta$ can be trisected using compass and straightedge if and only if the polynomial
\[
4x^3 - 3x - \cos(\theta)
\]
is reducible over $\mathbb{Q}(\cos(\theta))$.

\textbf{Bonus.} Choose your favorite trisectable angle $\theta$, and describe a step-by-step construction that performs the trisection; saying “I start with $\theta/3$ then triple it to make $\theta$ then I already have $\theta/3$ so I'm done” is cheating.
\end{problem}

\begin{problem}{4}
To $5$-sect an angle means to divide it by $5$.
\begin{enumerate}
\item Prove that the angles $2\pi$, $\pi$, $2\pi/3$, and $\pi/2$ can be $5$-sected by compass and straightedge.
\item Prove that a general angle cannot be $5$-sected by compass and straightedge.
\end{enumerate}
\end{problem}

\begin{problem}{5}
Let $p$ be an odd prime.
\begin{enumerate}
\item Prove that both $\mathbb{Q}(\zeta_p)$ and $\mathbb{Q}(\zeta_{2p})$ have degree $p-1$ over $\mathbb{Q}$.
\item Prove that both $\mathbb{Q}(\cos(2\pi/p))$ and $\mathbb{Q}(\cos(\pi/p))$ have degree $(p-1)/2$ over $\mathbb{Q}$.
\item Find the minimal polynomial of $\cos(2\pi/9)$ over $\mathbb{Q}$ and prove that it splits completely over $\mathbb{Q}(\cos(2\pi/9))$.
\end{enumerate}
\end{problem}

\begin{problem}{6}
For $3 \le n < 17$, determine whether a regular $n$-gon can be constructed by compass and straightedge. 
\end{problem}

\end{document}